### 2014-03-17 msc-thesis/chapters/3/sec-iced.tex
MSC final melanie edits.

My interview participants came from a large breadth of both open and closed source software development companies
and projects, using both  agile and waterfall based methodology, and from a wide spectrum of organizations, as shown in Table~\ref{tab:demo}.
My participants were invited based on their direct involvement in the actual writing of software for
their respective companies or projects. These participants' software development experience ranged from 3-25 years of experience
with an average of 8 years of experience.
In addition to software development, some participants were also invited based on their experience with project management
at some capacity. See Table~\ref{tab:demo} for more demographic details.

@@

While each of the 10 topics had a number of starter questions, interviews 
largely became discussions of developer experience and opinion as opposed to direct answers to any specific question.
However, not all participants had strong opinions or any experience on every category mentioned. For these participants, answers 
to the specific categories were not required or pressed upon. I attribute any non answer by a participant to
either lack of knowledge in their current project pertaining to the category or lack of experience in terms of
being apart of any one software project for extended periods of time.

@@

\subsubsection{questionnaire Procedure}

questionnaire participants were invited to participate in the questionnaire by email. No reminder email was sent as the
questionnaire responses were not connected with the invitation email addresses and thus participants who did respond
could not be identified. I directly emailed 1300 participants and ended with 78 responses
giving a response rate of 6\%. I attribute the low response rate with: the questionnaires
were conducted during the months of July and August while many participants may be away from their regular positions.
and my GitHub and Apache participants could not be verified as to whether or not they actively support the
email addresses used in the invitations. In addition, the questionnaire was considered by some to be long and require
more development experience than may have been typical of some of those invited to participate.

@@

As per my interview participants, I attempted to cover as great a breadth as possible of software developers in their practices, companies,
project, languages, etc. However, it would be extremely difficult to conduct a completely thorough study of indirect conflicts in practice
as there exist too many demographic options for participants. This same applies for questionnaire participants. This being said, questionnaire 
or interview results could become heavily biased towards a developers good or bad experiences with any given project.

To counter act this, my confirmatory interviews found high agreement for several of my findings. Therefore, I believe that
I have uncovered valuable insight regarding indirect conflicts in practice.

@@

My result analysis has indicated that: developers do not want awareness mechanisms which provide non actionable results; developers
would rather focus on curing existing problems rather than preventing potential issues;
and that there exists a gap in software evolution analytical tools between what is available and what practitioners need.

>>>>

My interview participants came from a large breadth of both open and closed source software development companies
and projects, using both  agile and waterfall based methodology, and from a wide spectrum of organizations, as shown in Table~\ref{tab:demo}.
My participants were invited based on their direct involvement in the actual writing of software for
their respective companies or projects. These participants' software development experience ranged from 3-25 years of experience
with an average of 8 years of experience.
In addition to software development, some participants were also invited based on their experience with project management. 
See Table~\ref{tab:demo} for more demographic details.

@@

While each of the 10 topics had a number of starter questions, interviews 
largely became discussions of developer experience and opinion as opposed to direct answers to any specific question.
However, not all participants had strong opinions or any experience on every category mentioned. For these participants, answers 
to the specific categories were not required or pressed upon. I attribute any non answer by a participant to
either lack of knowledge in their current project pertaining to the category or lack of experience in terms of
being a part of any one software project for extended periods of time.

@@

\subsubsection{Questionnaire Procedure}

Questionnaire participants were invited to participate in the questionnaire by email. No reminder email was sent as the
questionnaire responses were not connected with the invitation email addresses and thus participants who did respond
could not be identified. I directly emailed 1300 participants and ended with 78 responses
giving a response rate of 6\%. I attribute the low response rate with: the questionnaires
were conducted during the months of July and August while many participants may be away from their regular positions.
Also, my GitHub and Apache participants could not be verified as to whether or not they actively support the
email addresses used in the invitations. In addition, the questionnaire was considered by some to be long and required
more development experience than may have been typical of some of those invited to participate.

@@

I attempted to cover as great a breadth as possible of software developers in their practices, companies,
project, languages, etc. However, it would be extremely difficult to conduct a completely thorough study of indirect conflicts in practice
as there exist too many demographic options for participants. This same applies for questionnaire participants. This being said, questionnaire 
or interview results could become heavily biased towards a developer's good or bad experiences with any given project.

To counter this, my confirmatory interviews found high agreement for several of my findings. Therefore, I believe that
I have uncovered valuable insight regarding indirect conflicts in practice.

@@

My result analysis has indicated that: developers do not want awareness mechanisms which provide non actionable results; developers
would rather focus on curing existing problems rather than preventing potential issues;
and there exists a gap in software evolution analytical tools between what is available and what practitioners need.

### 2014-03-17 msc-thesis/chapters/4/chapter-discussion.tex
MSC final melanie edits.

As we have seen in Study 1, developers which are tied to source code objects can become a focal point of
indirect conflicts through the life of a project. For instance we can see from Table~\ref{tab:ratio} that
developers Daniel and Anne on Hibernate-ORM are always linked indirect conflicts when dependencies between
their source code objects change. This is an important observation to make when moving forward with indirect
conflicts.

@@

\textbf{Unwanted Awareness} \textit{Developers tend to only care about the awareness of other's activities, if it causes
a negative influence on their own work. Developers only want to hear about another developer's actions if it forces them
to take some action because of it. This limited awareness is quite different that what literature suggests, which is larger
awareness about most, if not all actions, and it also suggests why developers see a high amount of what they believe to be
false positives, as the changes being reported are not causing them to take action. These reported non actionable responses
lead to information overload and false positives.}

@@

The lack of prevention process and tools being used is believed to be due to 2 factors. The first being the identification of dependencies.
Even with an experienced system architect, identifying dependencies and notifying those involved is a daunting task 
which is ignored more than dealt with. The second being the knowledge of when a dependency will fail, also requires vast knowledge
of the product, more than anyone may have. This is compounded by the unique and sporadic nature of indirect conflicts. These 2 factors
are what ultimately have led to the amount of false positives and information overload seen in previous tools. The
abundance of detection and curative process and tools on the other hand shows once again developers willingness to debug
real issues, that maybe even could have been prevented, rather than prevent the issues in the first place. With
this lack of prevention and abundance of curative measures, the question ultimately presented in this area is if 
curative measures are more productive than preventative measures.

Dromey~\cite{Dromey:2003} has raised the debate of prevention versus cure in software development and how difficult a problem it is
to measure. The pros and cons of prevention versus cure I have identified, are similar to those of unwanted awareness,
in that they result in a trade off 
of where time is spent and how productive each side of the argument truly is. If prevention can be shown to be more effective in that
it reduces the number of indirect conflicts or time spent debugging them compared to the time taken to prevent them, should we then
not be moving developers into a more preventative mindset. If curative can be shown to be more effective in that it takes less time
to fix real issues compare to preventing potential ones, should we not be putting more time into automatic debugging, such as
Zeller's Delta Debugging~\cite{Zeller:2002:ICC} or program slicing techniques~\cite{Weiser:1982:PUS}, or automatic test case creation.

@@

This prevention versus cure discussion resides purely on the developer level and should be noted that it may not apply to system
designers, architects, or managerial stakeholders. 
From the project manager's interviewed it was shown that they are heavily favored towards planning and 
prevention (even though their prevention may be on more abstract levels than developers actually need) while leaving the curative
approaches to their developers. The prevention versus cure debate may have different outcomes depending on what level of abstraction
is being viewed in research.

@@

Indirect conflicts are more likely to occur,
from developer opinion, before a mature point is reached in the project's evolution. ``Once you get the API stable, people are 
better at communicating changes in regards to dependency concerns.''(This mature point may stem from a major release,
end of an iteration, a new feature being released or any point considered stable and reliable) However, this context of a 
mature point, as well as any other
potential contextual attributes, are rarely identified or accounted for in regards to indirect conflicts. Developers have said
that different change types may occur at different rates throughout a project's life time and that this
may drastically effect the outcomes of indirect conflict tools or processes.

@@

The direct implication this research has on tool creation is that of current adoption among developers. As was stated,
developers are more keen to invest their time at the detection and cure / resolution stages of indirect conflicts. That being 
said, focus at these two stages for tool developer will lead to larger adoption among developers.
It should be noted here that detection must come with an almost zero rate of false positives as
current tools (unit and integration and user testing), while they may not have 100\% recall, have almost 100\% precision.

>>>>

As we have seen in Study 1, developers that are tied to source code objects can become a focal point of
indirect conflicts through the life of a project. For instance we can see from Table~\ref{tab:ratio} that
developers Daniel and Anne on Hibernate-ORM are always linked indirect conflicts when dependencies between
their source code objects change. This is an important observation to make when moving forward with indirect
conflicts.

@@

\textbf{Unwanted Awareness} \textit{Developers tend to only care about the awareness of other's activities, if it causes
a negative influence on their own work. Developers only want to hear about another developer's actions if it forces them
to take some action because of it. This limited awareness is quite different from what literature suggests, which is larger
awareness about most, if not all actions, and it also suggests why developers see a high amount of what they believe to be
false positives, as the changes being reported are not causing them to take action. These reported non-actionable responses
lead to information overload and false positives.}

@@

The lack of prevention process and tools being used is believed to be due to 2 factors. The first being the identification of dependencies.
Even with an experienced system architect, identifying dependencies and notifying those involved is a daunting task 
which is ignored more than dealt with. The second being the knowledge of when a dependency will fail, also requires vast knowledge
of the product, more than anyone may have. This is compounded by the unique and sporadic nature of indirect conflicts. These 2 factors
are what ultimately have led to the amount of false positives and information overload seen in previous tools. The
abundance of detection and curative process and tools on the other hand shows once again developer's willingness to debug
real issues, that maybe even could have been prevented, rather than prevent the issues in the first place. With
this lack of prevention and abundance of curative measures, the question ultimately presented in this area is if 
curative measures are more productive than preventative measures.

Dromey~\cite{Dromey:2003} has raised the debate of prevention versus cure in software development and how difficult a problem it is
to measure. The pros and cons of prevention versus cure I have identified, are similar to those of unwanted awareness,
in that they result in a trade off 
of where time is spent and how productive each side of the argument truly is. If prevention can be shown to be more effective in that
it reduces the number of indirect conflicts or time spent debugging them compared to the time taken to prevent them, should we then
not be moving developers into a more preventative mindset; if curative can be shown to be more effective in that it takes less time
to fix real issues compare to preventing potential ones, should we not be putting more time into automatic debugging, such as
Zeller's Delta Debugging~\cite{Zeller:2002:ICC} or program slicing techniques~\cite{Weiser:1982:PUS}, or automatic test case creation.

@@

This prevention versus cure discussion resides purely on the developer level and should be noted that it may not apply to system
designers, architects, or managerial stakeholders. 
From the project managers interviewed it was shown that they heavily favored planning and 
prevention (even though their prevention may be on more abstract levels than developers actually need) while leaving the curative
approaches to their developers. The prevention versus cure debate may have different outcomes depending on what level of abstraction
is being viewed in research.

@@

Indirect conflicts are more likely to occur,
from developer opinion, before a mature point is reached in the project's evolution. ``Once you get the API stable, people are 
better at communicating changes in regards to dependency concerns.''(This mature point may stem from a major release,
end of an iteration, a new feature being released or any point considered stable and reliable.) However, this context of a 
mature point, as well as any other
potential contextual attributes, are rarely identified or accounted for in regards to indirect conflicts. Developers have said
that different change types may occur at different rates throughout a project's life time and that this
may drastically effect the outcomes of indirect conflict tools or processes.

@@

The direct implication this research has on tool creation is that of current adoption among developers. As was stated,
developers are more keen to invest their time at the detection and cure / resolution stages of indirect conflicts. That being 
said, focus at these two stages for tool development will lead to larger adoption among developers.
It should be noted here that detection must come with an almost zero rate of false positives as
current tools (unit and integration and user testing), while they may not have 100\% recall, have almost 100\% precision.

### 2014-03-17 msc-thesis/chapters/5/chapter-conclusion.tex
MSC final melanie edits.

While the techniques and tools for direct conflicts have had large success, tools either proposed or studied for 
indirect conflicts have had common issues of information overload, false positives, scalability, information distribution 
and many others. These issues has ultimately lead to poor developer interest in indirect conflict tools as well
as a failed adoption rate and partial failures of research.

To better understand these issues, this dissertation has focused on exploring the world of indirect 
conflicts through 4 studies. The first two studies focused on motivational circumstances which occur during 
the software development life cycle and cause indirect conflicts. Developers interactions were studied in order to create 
a tool which can aid in the work-flows around indirect conflicts. The second two studies presented a deeper investigation 
as to why most indirect conflict tools fail to attract developer interest through exploring the root causes of indirect conflicts 
and how tools should be properly built to support developer work flows.

@@

Technical dependencies are often used to predict software failures
in large software system~\cite{Pinzger:2008:DNP, Zimmermann:2008:PDU, Kim:2006:AIB}. However, human interactions as tied
to software objects can also be used to predict software failures.
Study 1 presented a method for detecting failure inducing pairs of developers inside
of technical networks based on code changes. The methodology used was to assigned developers to the source code artifacts
which they authored. Next, dependencies were identified between source code artifacts, implying dependencies between
developers. Lastly, failure inducing dependencies between developers were identified through using methods to identify
bugs in source code throughout the lifetime of a project. These developer pairs were used in the prediction
of future bugs as well as provide coordination recommendations for developers within a project. These bugs being predicted
directly correlate to indirect conflicts. Since the human dependencies being measured in Study 1 are directly tied to
software objects, we can say that the indirect conflicts being studied here are between developers themselves, while
the dependencies between the developers are found in the source code.

@@

Through the analysis of Study 1 as seen in Chapter 4, an important component of indirect conflicts are the developers
themselves and how their notions of other's work is perceived across a project. We can see from Study 1 results that
a developer Daniel may believe a source code artifact behaves in a particular way (from documentation or other forms
of communication for the artifact) so when that artifact changes its behavior, Daniel can become negatively impacted.
A human factor is present here as the person who makes the artifact change, say Anne, may not have communicated the
change to all other developers affected by the change. This is an important understanding that lead directly into the
creation of Impact (Study 2) as a way to mitigate the negative affects across human developers as a result of 
source code changes.

@@

Indirect conflicts are a significant issue with real world development, however, many proposed techniques and tools to mitigate
losses in this realm have been unsuccessful in attracting major support from developers (as was seen in Study 2). 
In Study 3, I conducted a qualitative study involving
19 interviewed developers from 12 organizations as well as 78 surveyed developers. Through these qualitative measures,
I provided characterization of indirect conflicts,
current developer work flow surrounding indirect conflicts, and what direction any future tools should follow in order to aid developers
in their work flow with indirect conflicts.

@@

As per current developer work-flows regarding indirect conflicts, I found that developers have 3 major mitigation strategies
to avoid indirect conflicts: ``Design by Contract'', add and deprecate, and personal experience. For catching indirect conflicts,
I found that use case coverage through proper testings (both unit and integration) are currently thought to be the best
developers can achieve. And finally, for the future of indirect conflict tools, developers made it clear that they would
prefer an improvement to resolution methods for indirect conflicts such as automatic or aided development techniques.

As per the analysis of Study 3, I have shown the disjoint of current awareness understanding versus the practical
awareness needs found in industry. This disjoint, caused by the difference of academic and practical understanding in awareness needs,
ultimately lead to tools with information overload and false positive issues. The debate of prevention versus cure was presented along with
the industrial tendency towards curative measures. While a curative approach may be favored by developers, further research is needed
to fully asses the positive and negative influences of prevention versus cure for productivity and quality concerns. Finally, I have shown
the gap of analytical evolution tools needed for dependency identification and indirect conflicts. This gap, unless addressed, may prevent
future industrial adoption of tools produced by researchers for lack of fit in industrial needs. This gap directly resulted in the final
study of this Dissertation to be completed (Study 4).

@@

Through the analysis in Chapter 4, I have shown the disjoint of current awareness understanding versus the practical
awareness needs found in industry. This disjoint, caused by the difference of academic and practical understanding in awareness needs,
ultimately lead to tools with information overload and false positive issues. This was quite evident from my tool Impact as
presented in Study 2. Even when the tool is based around the statistical results of Stud 1, the disjoint between what academics
determin developers want versus what they themselves want is quite obvious. The debate of prevention versus cure has been presented along with
the industrial tendency towards curative measures. However, while a curative approach may be favored by developers, further research is needed
to fully asses the positive and negative influences of prevention versus cure for productivity and quality concerns. Finally, I have shown
the gap of analytical evolution tools needed for dependency identification and indirect conflicts. This gap, unless addressed, may prevent
future industrial adoption of tools produced by researchers for lack of fit in industrial needs. While, I have shown a potential solution to address
a sub problem of the analysis gap through contextual patterns of indirect conflicts, many sub problems of the analysis gap remian problems
for future research.

>>>>

While the techniques and tools for direct conflicts have had large success, tools either proposed or studied for 
indirect conflicts have had common issues of information overload, false positives, scalability, information distribution 
and many others. These issues have ultimately led to poor developer interest in indirect conflict tools as well
as a failed adoption rate and partial failures of research.

To better understand these issues, this dissertation has focused on exploring the world of indirect 
conflicts through 4 studies. The first two studies focused on motivational circumstances which occur during 
the software development life cycle and cause indirect conflicts. Developer's interactions were studied in order to create 
a tool which can aid in the work-flows around indirect conflicts. The second two studies presented a deeper investigation 
as to why most indirect conflict tools fail to attract developer interest through exploring the root causes of indirect conflicts 
and how tools should be properly built to support developer work flows.

@@

Technical dependencies are often used to predict software failures
in large software system~\cite{Pinzger:2008:DNP, Zimmermann:2008:PDU, Kim:2006:AIB}. However, human interactions as tied
to software objects can also be used to predict software failures.
Study 1 presented a method for detecting failure inducing pairs of developers inside
of technical networks based on code changes. The methodology used was to assigned developers to the source code artifacts
which they authored. Next, dependencies were identified between source code artifacts, implying dependencies between
developers. Lastly, failure inducing dependencies between developers were identified through using methods to identify
bugs in source code throughout the lifetime of a project. These developer pairs were used in the prediction
of future bugs and to provide coordination recommendations for developers within a project. These bugs being predicted
directly correlate to indirect conflicts. Since the human dependencies being measured in Study 1 are directly tied to
software objects, we can say that the indirect conflicts being studied here are between developers themselves, while
the dependencies between the developers are found in the source code.

@@

Through the analysis of Study 1 as seen in Chapter 4, an important component of indirect conflicts are the developers
themselves and how their notions of other's work is perceived across a project. We can see from Study 1 results that
a developer Daniel may believe a source code artifact behaves in a particular way (from documentation or other forms
of communication for the artifact) so when that artifact changes its behavior, Daniel can become negatively impacted.
A human factor is present here as the person who makes the artifact change, say Anne, may not have communicated the
change to all other developers affected by the change. This is an important understanding that led directly into the
creation of Impact (Study 2) as a way to mitigate the negative affects across human developers as a result of 
source code changes.

@@

Indirect conflicts are a significant issue with real world development, however, many proposed techniques and tools to mitigate
losses in this realm have been unsuccessful in attracting major support from developers (as was seen in Study 2). 
In Study 3, I conducted a qualitative study involving
19 interviewed developers from 12 organizations as well as 78 surveyed developers. I provided characterization of indirect conflicts,
current developer work flow surrounding indirect conflicts, and what direction any future tools should follow in order to aid developers
in their work flow with indirect conflicts.

@@

As per current developer work-flows regarding indirect conflicts, I found that developers have 3 major mitigation strategies
to avoid indirect conflicts: ``Design by Contract'', add and deprecate, and personal experience. For catching indirect conflicts,
I found that use case coverage through proper testing (both unit and integration) are currently thought to be the best
developers can achieve. And finally, for the future of indirect conflict tools, developers made it clear that they would
prefer an improvement to resolution methods for indirect conflicts such as automatic or aided development techniques.

As per the analysis of Study 3, I have shown the disjoint of current awareness understanding versus the practical
awareness needs found in industry. This disjoint, caused by the difference of academic and practical understanding in awareness needs,
ultimately lead to tools with information overload and false positive issues. The debate of prevention versus cure was presented along with
the industrial tendency towards curative measures. While a curative approach may be favored by developers, further research is needed
to fully assess the positive and negative influences of prevention versus cure for productivity and quality concerns. Finally, I have shown
the gap of analytical evolution tools needed for dependency identification and indirect conflicts. This gap, unless addressed, may prevent
future industrial adoption of tools produced by researchers for lack of fit in industrial needs. This gap directly resulted in the final
study of this Dissertation to be completed (Study 4).

@@

Through the analysis in Chapter 4, I have shown the disjoint of current awareness understanding versus the practical
awareness needs found in industry. This disjoint, caused by the difference of academic and practical understanding in awareness needs,
ultimately led to tools with information overload and false positive issues. This was quite evident from my tool Impact as
presented in Study 2. Even when the tool is based around the statistical results of Study 1, the disjoint between what academics
determine developers want versus what they themselves want is quite obvious. The debate of prevention versus cure has been presented along with
the industrial tendency towards curative measures. However, while a curative approach may be favored by developers, further research is needed
to fully assess the positive and negative influences of prevention versus cure for productivity and quality concerns. Finally, I have shown
the gap of analytical evolution tools needed for dependency identification and indirect conflicts. This gap, unless addressed, may prevent
future industrial adoption of tools produced by researchers for lack of fit in industrial needs. While, I have shown a potential solution to address
a fraction of the analysis gap through contextual patterns of indirect conflicts, many other problems of the analysis gap remain problems
for future research.

### 2014-03-17 msc-thesis/frontmatter/abs.tex
MSC final melanie edits.

Awareness techniques have been proposed and studied to aid developer
understanding, efficiency, and quality of software produced. Some of these techniques have focused 
on either \textit{direct} or 
\textit{indirect conflicts} in order to prevent, detect, or resolve these conflicts as they arise
from a result of source code changes. While the techniques and tools for direct conflicts have had
large success, tools either proposed or studied for indirect conflicts have had common issues of
information overload, false positives, scalability, information distribution and many others. 
To better understand these issues, this dissertation will focus on exploring the world of indirect conflicts
through 4 studies. The first two studies presented will focus on motivational circumstances which occur
during the software development life cycle and cause indirect conflicts. Developers interactions are
studied in order to create a tool which can aid in the workflows around indirect conflicts. The second
two studies present a deeper investigation into why most indirect conflict tools fail to attrack developer
inerest through exploring the root causes of indirect conflicts and how tools should be properly built
to support developer workflows.

>>>>

Awareness techniques have been proposed and studied to aid developer
understanding, efficiency, and quality of software produced. Some of these techniques have focused 
on either \textit{direct} or 
\textit{indirect conflicts} in order to prevent, detect, or resolve these conflicts as they arise
from a result of source code changes. While the techniques and tools for direct conflicts have had
large success, tools either proposed or studied for indirect conflicts have had common issues of
information overload, false positives, scalability, information distribution and many others. 
To better understand these issues, this dissertation will focus on exploring the world of indirect conflicts
through 4 studies. The first two studies presented will focus on motivational circumstances which occur
during the software development life cycle and cause indirect conflicts. Developers interactions are
studied in order to create a tool which can aid in the workflows around indirect conflicts. The second
two studies present a deeper investigation into why most indirect conflict tools fail to attrack developer
interest through exploring the root causes of indirect conflicts and how tools should be properly built
to support developer workflows.
